### 2017-05-05 introduction.tex
need to cut a half page

The goal of this paper, an effective metric for computing distance between program executions (in particular, failing test cases), is motivated in large part by progress in automated testing.  Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{We claim testing/debugging complex systems software is well represented by compiler testing as an instance.}.  Unfortunately, the blessing of many failing test cases can become a curse for a developer.  Simply discovering which test cases in a large set of failures represent distinct faults (or already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.  For simple crashes and assertion violations, bucketing failures is sometimes easy.  When a failure is detected by differential testing \cite{Differential}, where two compilers or optimization levels disagree, the problem is so hard that it may turn potential users away from automated test generation \cite{PLDI13}.  Even once a fault is identified, fixing the problem can be very difficult.  Due to complexity of intermediate representation, interaction of optimizations, subtle language semantics, and other issues, debugging compiler faults may often be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
We propose the use of program mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.  The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault. Repairs can also provide  information for understanding and debugging discovered faults. The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.  Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).

@@

The key insight of this paper is a hypothesis with numerous possible applications, including a universal fuzzer taming metric that can also provide assistance in debugging.

\begin{quote}
{\bf Hypothesis 1:} If two test cases that fail can be made to succeed by the same syntactic modification to the original program $P$, this provides evidence that they fail due to the same fault.
\end{quote}

The core idea is hardly controversial:  most developers determine if they have fixed all faults in a set of failing test cases by re-running the tests after they fix at least one fault.  As a method for fuzzer taming or debugging assistance, however, this is useless.  Fortunately, developer-provided patches are not the only source of syntactic changes to a program.  Mutation analysis \cite{demillo1978hints,budd1980theoretical} applies large numbers of syntactic changes to a program, usually to evaluate a test suite \cite{mutant,justmutants}.  Some recent work has also considered these changes as a source of fixes or localizations for program faults \cite{achour,multilingual,MUSE,DebroyMutant}.  Our experience, as well as empirical data \cite{GopinathMutants}, suggests that very few compiler faults are due to the kind of simple syntactic changes found in mutants.  However, this does not matter for {\bf Hypothesis 1};  what matters is that failures, which correspond to faults, can be made to succeed.  When a test case $t$ fails for program $P$ but succeeds for mutant $m$ of $P$ we say that $m$ \emph{repairs} $t$.  In most cases this does not mean that $m$ is the inverse of a fault, but that $m$ somehow avoids the consequence of a fault.  It is not important that repairs actually fix anything, if all we seek is a way to decide whether failures are due to the same fault.

Hypothesis 1 is in fact a consequence of a more general (and difficult to investigate) hypothesis.  What is interesting about a program execution?  For many purposes, the only interesting properties of an execution are those that in some way relate to the correctness properties of the program, assuming we have a good set of correctness properties.  These may include assertions, differential testing \cite{Differential,ICSEDiff}, timing constraints, temporal logic formulas, or any other checkable properties.  What, then, makes two executions similar?  Our proposal is to say that:

\begin{quote}
{\bf Hypothesis 2:} Two executions of $P$ are similar to the degree that changes to $P$ produce the same changes in correctness of the executions.
\end{quote}

This leads to the idea of comparing executions by their response to mutants.   In the presence of multiple correctness properties, any two executions can be compared, not just two failures, since a mutant can cause a failure to violate additional or different properties.
%Where there are multiple, independent, correctness properties, this even allows us to compare failing and successful executions, since a mutant can cause a failure to violate a different property.
{\bf Hypothesis 2} at first may suggest that this idea subsumes past notions of similarity, such as spectra of code coverage, since if two executions do not cover a statement, they both cannot alter correctness if that statement is altered.  The metrics are in fact simply incomparable:  for example, differences in coverage that never have an effect on correctness cannot contribute to similarity in our approach; with weak correctness properties most executions are very similar.  This is similar to the idea of \emph{checked coverage}:  code that cannot impact correctness is not interesting \cite{CheckCov}.  Hypothesis 2 also suggests an additional hypothesis related to hypothesis 1.  This is {\bf Hypothesis 1a:} If two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.

@@

These hypothesis do not claim that being repaired by the same mutant perfectly predicts fault equivalence.  To understand why that seems highly unlikely, consider why a mutant of a compiler might cause a test case to stop failing.  Many, likely most, compiler faults are due to incorrect optimizations.  Turning an optimization off can cause a program to compile correctly.  Optimizations are generally guarded by conditional switches --- in some cases, so that the optimization can be turned off by a user, in other cases when a property of the code in question prevents the optimization.  If two failures become successes due to turning off an optimization, it is certainly some level of evidence they are the same fault, compared to the baseline probability.  Intuitively, the evidence provided seems much stronger than the fact that, e.g., the two test cases execute a shared line of code \cite{RepsSpectra}.  However, optimization guards range from guarding very specific fragments of code to guarding entire modules with many functions and related sub-optimizations.  If a mutant prevents a very specific optimization (consisting of a few lines of code) from executing, and repairs two tests, that is powerful evidence they are related.  If a mutant turns off a large number of optimizations, it is weak evidence.  Must we then characterize mutant granularity?  No.

@@

@@

The contributions of this paper are: hypotheses about how to measure distance between program executions, and practical proposals for using such a metric in fuzzer taming and in novel fault localization and explanation methods.  Evaluating the effectiveness of these methods also provides experimental evidence as to the validity of our hypotheses.  Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over previous results \cite{PLDI13}, in multiple ways.  Our preliminary localization approaches compare favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual}, for the mutation analysis budget in our experiments.

>>>>

The proposed techniques in this paper are motivated in large part by progress in automated testing.  Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in modern optimizing compilers.  Unfortunately, the blessing of many failing test cases can become a curse for a developer.  Simply discovering which test cases in a large set of failures represent distinct faults (or already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.  For simple crashes and assertion violations, bucketing failures is sometimes easy.  When a failure is detected by differential testing \cite{Differential}, where two compilers or optimization levels disagree, the problem is so hard that it may turn potential users away from automated test generation \cite{PLDI13}.  Even once a fault is identified, fixing the problem can be very difficult.  Due to complexity of intermediate representation, interaction of optimizations, subtle language semantics, and other issues, debugging compiler faults may often be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
We propose the use of program mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.  The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault. Repairs can also provide  information for understanding and debugging discovered faults. \comment{The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.  Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).}

@@

The key insight of this paper is based on an idea with numerous possible applications: \emph{if two test cases that fail can be made to succeed by the same small modification to the original program $P$, this provides evidence that they fail due to the same fault}.  The idea is hardly controversial:  most developers determine if they have fixed all faults in a set of failing test cases by re-running the tests after they fix at least one fault.  As a method for fuzzer taming or debugging assistance, however, this is useless.  Fortunately, developer-provided patches are not the only source of changes to a program.  Mutation analysis \cite{demillo1978hints,budd1980theoretical} applies large numbers of small syntactic changes to a program, usually to evaluate a test suite \cite{mutant,justmutants}.  Some recent work has also considered these changes as a source of fixes or localizations for program faults \cite{achour,multilingual,MUSE,DebroyMutant}.  Our experience, as well as empirical data \cite{GopinathMutants}, suggests that very few compiler faults are due to the kind of simple syntactic changes found in mutants.  However, what should matter is that failures, which correspond to faults, can be made to succeed.  When a test case $t$ fails for program $P$ but succeeds for mutant $m$ of $P$ we say that $m$ \emph{repairs} $t$.  In most cases this does not mean that $m$ is the inverse of a fault, but that $m$ somehow avoids the consequence of a fault.  It is not important that repairs actually fix anything, if all we seek is a way to decide whether failures are due to the same fault.

This idea is in fact a consequence of a more general (and difficult to investigate) hypothesis.  What is interesting about a program execution?  For many purposes, the only interesting properties of an execution are those that in some way relate to the correctness properties of the program, assuming we have a good set of correctness properties.  These may include assertions, differential testing \cite{Differential,ICSEDiff}, timing constraints, temporal logic formulas, or any other checkable properties.  What, then, makes two executions similar?  Our proposal is that \emph{two executions of $P$ are similar to the degree that changes to $P$ produce the same changes in correctness of the executions.}

This leads to the idea of comparing executions by their response to mutants.   In the presence of multiple correctness properties, in principle any two executions can be compared, not just two failures, since a mutant can cause a failure to violate additional or different properties.   However, for practical purposes the most interesting implication is that if two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.

\comment{
%Where there are multiple, independent, correctness properties, this even allows us to compare failing and successful executions, since a mutant can cause a failure to violate a different property.
{\bf Hypothesis 2} at first may suggest that this idea subsumes past notions of similarity, such as spectra of code coverage, since if two executions do not cover a statement, they both cannot alter correctness if that statement is altered.  The metrics are in fact simply incomparable:  for example, differences in coverage that never have an effect on correctness cannot contribute to similarity in our approach; with weak correctness properties most executions are very similar.  This is similar to the idea of \emph{checked coverage}:  code that cannot impact correctness is not interesting \cite{CheckCov}.  Hypothesis 2 also suggests an additional hypothesis related to hypothesis 1.  This is {\bf Hypothesis 1a:} If two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.
}

@@

This not a claim that being repaired by the same mutant perfectly predicts fault equivalence.  To understand why that seems highly unlikely, consider why a mutant of a compiler might cause a test case to stop failing.  Many, likely most, compiler faults are due to incorrect optimizations.  Turning an optimization off can cause a program to compile correctly.  Optimizations are generally guarded by conditional switches --- in some cases, so that the optimization can be turned off by a user, in other cases when a property of the code in question prevents the optimization.  If two failures become successes due to turning off an optimization, it is presumably \emph{some} level of evidence they are the same fault, compared to the baseline probability.  Intuitively, the evidence provided seems much stronger than the fact that, e.g., the two test cases execute a shared line of code \cite{RepsSpectra}.  However, optimization guards range from guarding very specific fragments of code to guarding entire modules with many functions and related sub-optimizations.  If a mutant prevents a very specific optimization (consisting of a few lines of code) from executing, and repairs two tests, that is powerful evidence they are related.  If a mutant turns off a large number of optimizations, it is weak evidence.  Must we then characterize mutant granularity?  No.

@@

This paper investigates the utility of using distance metrics based on the causal approach suggested above.  We show that such a metric can improve, in a statistically significant way, the ability to distinguish tests based on which faults they trigger.  Furthermore, fault localization based on the same metrics can successfully pinpoint the location of some compiler faults sufficiently to be likely to substantially aid debugging, and, when applicable, performs very well for some faults studied in recent papers on mutant-based fault localization.

@@

The contributions of this paper include our ideas about how to measure distance between program executions, but more importantly some practical proposals for using such a metric in compiler fuzzer taming and fault localization.  Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over best previous results \cite{PLDI13}, in multiple ways.  Our best preliminary localization approach also compares favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual,Metallaxis}, albeit with a more limited applicability (but a lower cost).

### 2017-05-05 localizations.tex
need to cut a half page

Automated fault localization \cite{FaultSurvey,Jones2002} provides users with
information about the likely location of a fault in a program's source
code.  Error explanation \cite{GroceError} also provides information
about the causal structure of a fault --- how it can be avoided, what
triggers it --- a kind of ``fault story.''  Parnin and Orso recently
suggested that fault localization methods are not actually providing a
great deal of assistance to real users in debugging \cite{AutoHelp}.
Automated fault localization has not been widely adopted in industry.

@@

By definition, when a mutant repairs a test case, it changes the
behavior of the test.  In some cases, this behavioral change is
limited to changes in program state and data values.  However, the
vast majority of mutant repairs also modify the code coverage of a
test; in our experiments, there were no repairs that did not change
coverage.  If a fault can be avoided by, for example, disabling a
problematic optimization in a compiler, this naturally leads to
changed coverage.  If the fault is due to a bad conditional, where a
program path should include additional code, the repair will often
change the execution to include the omitted code.  As a localization
for a fault, then, we can examine the source statements whose
  coverage changed most often in repairs.  Early experiments with this
  approach showed that a useful coverage change localization must  account for the frequency of a
change.  Some coverage changes are extremely common across all
failures, faults, and repairing mutants,
and therefore likely not related to any specific fault.

@@

That is, the ranking is based on the number of times a statement's
coverage changes its value (covered or not covered) from the original
execution of $f$ in executions of mutants that repair $f$.  Changes are weighted by their inverse frequency over all
failures and all mutants.
Higher values for $r$ result in a higher ranking (more likely to
be faulty).  
%Computing $U$ accounts for the fact that in our
%experiments, a few statements were added or removed for every single
%r

The repairing mutants themselves can also be used for localization and
explanation.  While few will be equivalent to actual fixes, they
will sometimes overlap the incorrect code, and will often be closely
related to incorrect code, semantically.  Since many failures are
fixed by a large number of mutants, it is important to rank the
mutants to produce a localization.  Based on the onion-ring model proposed above, this is simple: each mutant is ranked as
a localization/explanation based on the \emph{most distant failure it also
  repairs.}  If a given mutant only repairs failures that have similar
repair vectors to the test case being localized, it will be highly
ranked.  If it repairs even one very dissimilar failure, then it will
be considered a poor localization.

@@

The coverage localization also provides a kind of explanation, though
a less direct one.  For any
interesting line of code in the coverage-change-based localization,
the user can investigate a particular instance of the change,
by selecting a single mutant and stepping through the execution of the
failing test case in a debugger to the point of change.  The user can also
browse through all mutants that produce this coverage change for the test
case.

Another advantage over many statistical approaches to
localization is that our methods are not potentially confused in
settings with multiple faults.  Each localization/explanation is based
on a single failing test case, which in most cases fails due to only one
fault.  In fact, other faults help both methods rank localizations.

>>>>

Automated fault localization \cite{FaultSurvey,Jones2002} provides users with
information about the likely location of a fault in a program's source
code.  \comment{Error explanation \cite{GroceError} also provides information
about the causal structure of a fault --- how it can be avoided, what
triggers it --- a kind of ``fault story.''}  Parnin and Orso recently
suggested that fault localization methods are not actually providing a
great deal of assistance to real users in debugging \cite{AutoHelp}.
Automated fault localization has not been widely adopted in industry.

@@

\comment{
By definition, when a mutant repairs a test case, it changes the
behavior of the test.  In some cases, this behavioral change is
limited to changes in program state and data values.  However, the
vast majority of mutant repairs also modify the code coverage of a
test; in our experiments, there were no repairs that did not change
coverage.  If a fault can be avoided by, for example, disabling a
problematic optimization in a compiler, this naturally leads to
changed coverage.  If the fault is due to a bad conditional, where a
program path should include additional code, the repair will often
change the execution to include the omitted code.  As a localization
for a fault, then, we can examine the source statements whose
  coverage changed most often in repairs.  Early experiments with this
  approach showed that a useful coverage change localization must  account for the frequency of a
change.  Some coverage changes are extremely common across all
failures, faults, and repairing mutants,
and therefore likely not related to any specific fault.

@@

That is, the ranking is based on the number of times a statement's
coverage changes its value (covered or not covered) from the original
execution of $f$ in executions of mutants that repair $f$.  Changes are weighted by their inverse frequency over all
failures and all mutants.
Higher values for $r$ result in a higher ranking (more likely to
be faulty).  
%Computing $U$ accounts for the fact that in our
%experiments, a few statements were added or removed for every single
%r
}
Repairing mutants can be used for localization and
explanation.  While few will be equivalent to actual fixes, they
will sometimes overlap the incorrect code, and will often be closely
related to incorrect code, semantically.  Since some tests are
repaired by a large number of mutants, it is important to rank the
mutants to produce a localization.  Based on the onion-ring model proposed above, this is simple: each mutant is ranked as
a localization/explanation based on the \emph{most distant failure it also
  repairs.}  If a given mutant only repairs failures that have similar
repair vectors to the test case being localized, it will be highly
ranked.  If it repairs even one very dissimilar failure, then it will
be considered a poor localization.

@@

An advantage of this method over some statistical approaches to
localization is that it is not potentially confused in
settings with multiple faults.  Each localization/explanation is based
on a single failing test case, which in most cases fails due to only one
fault.  In fact, other faults (and their failures) theoretically \emph{help} {\bf Repair} rank the mutants.

In general, we expect {\bf Repair} to be used in the context of
compiler (or other complex system) fuzzing, where there are numerous
failing tests, likely from multiple faults.  In such cases, the $f$ to
be localized is presumably selected by FPF.  However, in case our
localization approach is used in a setting where single faults are
expected, we propose an alternative method: let $f$ be an arbitrary
test with as few repairing mutants as possible.

\comment{
The coverage localization also provides a kind of explanation, though
a less direct one.  For any
interesting line of code in the coverage-change-based localization,
the user can investigate a particular instance of the change,
by selecting a single mutant and stepping through the execution of the
failing test case in a debugger to the point of change.  The user can also
browse through all mutants that produce this coverage change for the test
case.

Another advantage over many statistical approaches to
localization is that our methods are not potentially confused in
settings with multiple faults.  Each localization/explanation is based
on a single failing test case, which in most cases fails due to only one
fault.  In fact, other faults help both methods rank localizations.}

### 2017-05-05 metrics.tex
need to cut a half page

For every failing test case, its response to mutants defines a bit-vector, with
length equal to the number of mutants that repaired any test case,
where a 1 bit indicates that the mutant in question repairs this test
case.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
by 100 mutants.  If these two test cases share 90 common repairing
mutants, and disagree for 10, they have a Hamming
distance of 20.  This makes them ``less similar'' than two test cases,
$t_1'$ and $t_2'$, where $t_1'$ is repaired by 5 mutants, $t_2'$ is
repaired by 5 mutants, and \emph{none of those mutants overlap.}   A Jaccard distance, also used (as similarity) in some fault
localization algorithms \cite{Pinpoint}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
%\vspace{-0.08in}
\[d(u,v) = 
\begin{cases}
\frac{\sum_{i=0}^{n-1} 1\ \text{if}\ u_i \neq v_i\ \text{else}\ 0}{\sum_{i=0}^{n-1} 1
\  \text{if}\ u_i \neq 0 \vee v_i \neq 0\ \text{else}\ 0} & \exists i:u_i\neq 0 \vee v_i\neq 0\\
0 & otherwise
\end{cases}
\]

This metric, the portion of mismatches over bits that are 1 in either
vector, matches our intuitions about the above comparisons\footnote{In Section \ref{conc} we discuss possible uses of $d$
requiring comparison of both failing and passing test cases:  in this 
case, we would use three-valued vectors, where 1 indicates a 
failing property succeeds and -1 that a succeeding property 
fails.}.  It
essentially formalizes the intuition of the onion-ring model, where
more matching repairs indicates a much higher probability two failures
are due to the same fault, even if not all repairs match.

>>>>

For every failing test case, its response to mutants defines a bit-vector, with
length equal to the number of mutants that repaired any test case,
where a 1 bit indicates that the mutant in question repairs this test
case\footnote{By \emph{repairing} a failing test we mean that a mutant
causes the test to pass, and the mutant does not cause
other, previously passing tests, to fail.}.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
by 100 mutants.  If these two test cases share 90 common repairing
mutants, and disagree for 10, they have a Hamming
distance of 20.  This makes them ``less similar'' than two test cases,
$t_1'$ and $t_2'$, where $t_1'$ is repaired by 5 mutants, $t_2'$ is
repaired by 5 mutants, and \emph{none of those mutants overlap.}   A Jaccard distance, also used (as similarity) in some fault
localization algorithms \cite{Pinpoint}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
%\vspace{-0.08in}
\[d(u,v) = 
\begin{cases}
\frac{\sum_{i=0}^{n-1} 1\ \text{if}\ u_i \neq v_i\ \text{else}\ 0}{\sum_{i=0}^{n-1} 1
\  \text{if}\ u_i \neq 0 \vee v_i \neq 0\ \text{else}\ 0} & \exists i:u_i\neq 0 \vee v_i\neq 0\\
0 & otherwise
\end{cases}
\]

This metric, the portion of mismatches over bits that are 1 in either
vector, matches our intuitions about the above comparisons\footnote{In Section \ref{conc} we discuss uses of $d$
requiring comparison of both failing and passing test cases:  in this 
case, we would use three-valued vectors, where 1 indicates a 
failing property succeeds and -1 that a succeeding property 
fails.}.  It
essentially formalizes the intuition of the onion-ring model, where
more matching repairs indicates a much higher probability two failures
are due to the same fault, even if not all repairs match.
